---
title: "4305 Script"
output: pdf_document
date: "2024-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4305 Datathon Instructions

The following document is a series of instructions on running the reproducible codes supplied by group UG15.

The main focus of the pricing task, including data preparation, modelling, and the pricing output, will be included in this document, under the section *Main Script*.

In addition to this, the submitted folder contains supplementary codes to run the other sections included throughout our modelling process; this includes EDA findings and the constructions of models other than our credibility adjusted EBM. The instructions to run these are as follows:

-   For files under the 'EDA' folder, these should each be run out of a new R session, and are self-contained.

-   For files under the 'Models' folder, the file 'Adding_data' from the 'Data Preparation' folder must be run first within the same R session, as this file creates the required training and test datasets.

## Main Script

### Data Preparation

The first step of our main process is creating first cleaning the data. Here, labels were adjusted and standardised, and earned data was grouped by exposure ID.

```{r}
### Load packages
library(ggplot2)
library(dplyr)
library(glmnet)
library(randomForest)
library(plotly)
library(caret)
library(tidyr)
library(stringdist)
library(cluster)
library(factoextra)


### Set Plot Themes
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5, size = 10, face = 'bold'))
theme_update(text = element_text(size = 10))

#---- Data Preparation----

earned <- read.csv('UNSW_earned_data_adjusted_Sep27.csv')

# Remove time part from UW_Date and nb_policy_first_inception_date
earned$UW_Date <- gsub(" .*", "", earned$UW_Date)
earned$nb_policy_first_inception_date <- gsub(" .*", "", earned$nb_policy_first_inception_date)

# Convert to Date class assuming the format is day/month/year (DD/MM/YYYY)
earned$UW_Date <- as.Date(earned$UW_Date, format = "%d/%m/%Y")
earned$nb_policy_first_inception_date <- as.Date(earned$nb_policy_first_inception_date, format = "%d/%m/%Y")
earned$person_dob <- as.Date(earned$person_dob, format = "%d/%m/%Y")

# Use gsub to remove everything in parentheses, including the parentheses themselves
earned$quote_time_group <- gsub("\\s*\\(.*?\\)", "", earned$quote_time_group)

# Remove the exposure_id_1 column using the subset function
earned <- subset(earned, select = -exposure_id_1)

# Replace age ranges with categories
earned$pet_de_sexed_age <- gsub("0-3 (mo|months)", "de_sex_0-3mo", earned$pet_de_sexed_age)
earned$pet_de_sexed_age <- gsub("4-6 mo", "de_sex_4-6mo", earned$pet_de_sexed_age)   # 4-6 months -> Young
earned$pet_de_sexed_age <- gsub("7-12 (mo|months)", "de_sex_7-12mo", earned$pet_de_sexed_age) # 7-12 months -> Juvenile
earned$pet_de_sexed_age <- gsub("1-2 yr", "de_sex_1-2yr", earned$pet_de_sexed_age)    # 1-2 years -> Adult
# Replace "2+ yr" with "Mature", escaping the "+" character
earned$pet_de_sexed_age <- gsub("2\\+ yr", "de_sex_2+yr", earned$pet_de_sexed_age)    # 2+ years -> Mature
earned$pet_de_sexed_age <- gsub("2\\+ years", "de_sex_2+yr", earned$pet_de_sexed_age) # In case of "2+ years" as well

# Replace blank entries with "Never de-sexed" and keep "Not Sure"
earned$pet_de_sexed_age <- ifelse(earned$pet_de_sexed_age == "", "Never de-sexed", earned$pet_de_sexed_age)
earned$pet_de_sexed_age <- ifelse(earned$pet_de_sexed_age == "Not Sure", "Not Sure", earned$pet_de_sexed_age)

# Convert the column to a factor (categorical variable)
earned$pet_de_sexed_age <- factor(earned$pet_de_sexed_age, 
                                  levels = c("de_sex_0-3mo", "de_sex_4-6mo", "de_sex_7-12mo", "de_sex_1-2yr", "de_sex_2+yr", "Not Sure", "Never de-sexed"))

# Encoding pet_gender, pet_de_sexed, nb_address_type_adj, is_multi_pet_plan, quote_time_group
earned$pet_gender <- factor(earned$pet_gender, levels = c("male", "female"))
earned$pet_de_sexed <- factor(earned$pet_de_sexed, levels = c("TRUE", "FALSE"))
earned$nb_address_type_adj <- factor(earned$nb_address_type_adj)
earned$is_multi_pet_plan <- factor(earned$is_multi_pet_plan, levels = c("TRUE", "FALSE"))
earned$quote_time_group <- factor(earned$quote_time_group)

# Removing the 'X' column
earned <- earned[ , !(names(earned) %in% c("X"))]
earned <- earned[ , !(names(earned) %in% c("row_num"))]

# Summarise the data, accounting for all columns
earned <- earned %>%
  group_by(exposure_id) %>%
  summarise(
    # Date Columns
    UW_Date = max(UW_Date),  # Most recent underwriting date
    nb_policy_first_inception_date = min(nb_policy_first_inception_date, na.rm = TRUE),  # Earliest inception date
    
    # Numeric Columns
    tenure = max(tenure, na.rm = TRUE),  # Maximum tenure
    pet_age_months = mean(pet_age_months, na.rm = TRUE),  # Average pet age in months
    pet_age_years = mean(pet_age_years, na.rm = TRUE),  # Average pet age in years
    nb_contribution = mean(nb_contribution, na.rm = TRUE), # Average contribution
    nb_contribution_excess = mean(nb_contribution_excess, na.rm = TRUE),  # Average contribution excess
    nb_excess = mean(nb_excess, na.rm = TRUE),  # Average excess
    nb_number_of_breeds = max(nb_number_of_breeds, na.rm = TRUE),  # Maximum number of breeds
    nb_average_breed_size = mean(nb_average_breed_size, na.rm = TRUE),  # Average breed size
    earned_units = sum(earned_units, na.rm = TRUE),  # Sum of earned units
    
    # Boolean Columns
    is_multi_pet_plan = any(is_multi_pet_plan == "TRUE"),  # TRUE if any multi-pet plan exists
    pet_de_sexed = first(pet_de_sexed),  # First occurrence of pet de-sexed
    
    # Categorical Columns
    pet_gender = first(pet_gender),  # First occurrence of pet gender
    pet_de_sexed_age = first(pet_de_sexed_age),  # First occurrence of pet de-sexed age
    nb_address_type_adj = first(nb_address_type_adj),  # First occurrence of address type
    nb_postcode = first(nb_postcode),  # First postcode
    nb_state = first(nb_state),  # First state
    nb_suburb = first(nb_suburb),  # First suburb
    nb_breed_type = first(nb_breed_type),  # First breed type
    nb_breed_trait = first(nb_breed_trait),  # First breed trait
    nb_breed_name_unique = first(nb_breed_name_unique),  # First breed name
    nb_breed_name_unique_concat = first(nb_breed_name_unique_concat),  # First concatenated breed name
    quote_time_group = first(quote_time_group),  # First quote time group
    
    # Person-related columns (if applicable)
    person_dob = first(person_dob),  # First person date of birth
    owner_age_years = max(owner_age_years, na.rm = TRUE),  # Average owner age
    
    # Additional Columns
    pet_is_switcher = first(pet_is_switcher),  # First occurrence of switcher
    lead_date_day = first(lead_date_day),  # First occurrence of lead date day
    quote_date = first(quote_date),  # First quote date
  )

# Identify columns with NAs and count the number of NAs in each
na_counts <- sapply(earned, function(x) sum(is.na(x)))
na_counts <- na_counts[na_counts > 0]  # Filter to show only columns with NAs

earned$person_dob

# Print the results
print(na_counts)


earned <- earned[earned$tenure >= 0, ]

# Remove rows with NA in person_dob
earned <- earned[!is.na(earned$person_dob), ]

claims <- read.csv('UNSW_claims_data.csv')

# Convert 'claim_start_date' to Date type
claims$claim_start_date <- as.Date(claims$claim_start_date, format="%d/%m/%Y")

# Ensure 'claim_paid' and 'total_claim_amount' are numeric
claims$claim_paid <- as.numeric(claims$claim_paid)
claims$total_claim_amount <- as.numeric(claims$total_claim_amount)

# Remove the 'tenure' column from the 'claims' dataset
claims <- claims[, !names(claims) %in% 'tenure']

#----Creating severity dataset----

# Remove rows where claim_paid == 0
claims <- claims %>%
  filter(claim_paid != 0)

#----Calculating Frequency of claims per exposure----

# Count the number of claims for each exposure_id and store it in 'claim_counts'
claim_counts <- claims %>%
  group_by(exposure_id) %>%
  summarise(claim_count = n()) %>%
  ungroup()

# Step 2: Merge the claim frequencies into the 'earned' dataset without modifying 'claims'
frequency_data <- earned %>%
  left_join(claim_counts, by = "exposure_id") %>%
  mutate(claim_count = ifelse(is.na(claim_count), 0, claim_count))


### Creating the test dataset for the overall RMSE calculation (test_data_prem)
# Step 1: Expand the 'claims' dataset so each row corresponds to a separate claim, with claim_count = 1
# For each exposure_id, create a row for each individual claim
expanded_claims <- claims %>%
  mutate(claim_count = 1, 
         claim_paid = ifelse(is.na(claim_paid), 0, claim_paid))  # Set claim_paid to 0 if it is NA

frequency_data2 <- earned %>%
  left_join(expanded_claims, by = "exposure_id") %>%
  mutate(claim_count = ifelse(is.na(claim_count), 0, claim_count),
         claim_paid = ifelse(is.na(claim_paid), 0, claim_paid))  # Set claim_paid to 0 for non-claim rows

```

Next, we imported data from the ABS and DogTime.

```{r}
#----Adding in sets of potential predictors

### Adding in SA2, SA3 and SA4 location

SA3 <- read.csv('POSTCODE_SA3_MAPPING.csv')

# Ensure unique mapping for each postcode in SA3, picking the first match
SA3_unique <- SA3 %>%
  group_by(POSTCODE) %>%
  summarise(SA3_CODE_2011 = first(SA3_CODE_2011))  # Pick the first match for SA3

# Perform the left join to add SA3_CODE_2011
frequency_data <- frequency_data %>%
  left_join(SA3_unique, by = c("nb_postcode" = "POSTCODE"))

#----Incorporating ABS related data----
income_sa3 <- read.csv('income_sa3.csv')

library(dplyr)
frequency_data <- frequency_data %>%
  left_join(income_sa3 %>% select(Code, Median_employee_income_SA3, Mean_employee_income_SA3), 
            by = c("SA3_CODE_2011" = "Code"))

added_var_abs <- read.csv('add_abs_variables.csv')
colnames(added_var_abs)

# Left join frequency_data with added_var_abs using SA3_CODE_2011 and selecting specific columns from added_var_abs
frequency_data <- frequency_data %>%
  left_join(
    added_var_abs %>% 
      select(Code, 
             Total_number_bus, 
             Number_non.employing_bus,  # Updated to the correct column name
             Total_employed_aged_15_and_over, 
             Median_price_established_house_transfers, 
             Value_private_sector_houses, 
             Agriculture_forestry_fishing_per, 
             Health_care_social_assistance_per, 
             Professional_scientific_technical_services_per, 
             Number_employee_jobs_agriculture_forestry_fishing, 
             Number_employee._jobs_health_care,  # Updated to the correct column name
             Total_persons_aged_15_years_over, 
             Persons_who_made_HELP_repayment, 
             Number_employee_jobs_professional_scientific_technical, 
             Graduate_diploma_perc, 
             Bachelor_degree_perc, 
             Completed_year_12_perc, 
             Worked_from_home, 
             Average_household_size, 
             Lone_person_households, 
             Median_weekly_household_rent, 
             Persons_who_made_gift_or_donation, 
             Total_protected_land_area, 
             National_parks, 
             Nature_reserves, 
             X_1.499_per_week, 
             Participation_rate, 
             Professionals_perc, 
             Bachelor_degree_perc.1, 
             X_3000_or_more_per_week, 
             Managers, 
             Unemployment_rate, 
             X_2000_2999_per_week, 
             Not_in_labour_force), 
    by = c("SA3_CODE_2011" = "Code")
  )

# Convert appropriate columns to factor or numeric
frequency_data <- frequency_data %>%
  mutate(
    # Convert logical columns to factor
    is_multi_pet_plan = as.factor(is_multi_pet_plan),
    pet_de_sexed = as.factor(pet_de_sexed),
    pet_gender = as.factor(pet_gender),
    pet_de_sexed_age = as.factor(pet_de_sexed_age),
    nb_address_type_adj = as.factor(nb_address_type_adj),
    nb_breed_type = as.factor(nb_breed_type),
    
    # Convert character columns to factor if needed
    nb_state = as.factor(nb_state),
    nb_suburb = as.factor(nb_suburb),
    
    # Ensure numeric variables are correctly encoded
    Median_employee_income_SA3 = as.numeric(as.character(Median_employee_income_SA3)),
    Mean_employee_income_SA3 = as.numeric(as.character(Mean_employee_income_SA3)),
    Total_number_bus = as.numeric(as.character(Total_number_bus)),
    Number_non.employing_bus = as.numeric(as.character(Number_non.employing_bus)),
    Total_employed_aged_15_and_over = as.numeric(as.character(Total_employed_aged_15_and_over)),
    Agriculture_forestry_fishing_per = as.numeric(as.character(Agriculture_forestry_fishing_per)),
    Health_care_social_assistance_per = as.numeric(as.character(Health_care_social_assistance_per)),
    Professional_scientific_technical_services_per = as.numeric(as.character(Professional_scientific_technical_services_per)),
    Number_employee_jobs_agriculture_forestry_fishing = as.numeric(as.character(Number_employee_jobs_agriculture_forestry_fishing)),
    Number_employee._jobs_health_care = as.numeric(as.character(Number_employee._jobs_health_care)),
    Worked_from_home = as.numeric(as.character(Worked_from_home)),
    Lone_person_households = as.numeric(as.character(Lone_person_households)),
    X_1.499_per_week = as.numeric(as.character(X_1.499_per_week)),
    Participation_rate = as.numeric(as.character(Participation_rate)),
    Professionals_perc = as.numeric(as.character(Professionals_perc)),
    Bachelor_degree_perc.1 = as.numeric(as.character(Bachelor_degree_perc.1)),
    X_3000_or_more_per_week = as.numeric(as.character(X_3000_or_more_per_week)),
    Managers = as.numeric(as.character(Managers)),
    Unemployment_rate = as.numeric(as.character(Unemployment_rate)),
    X_2000_2999_per_week = as.numeric(as.character(X_2000_2999_per_week)),
    Not_in_labour_force = as.numeric(as.character(Not_in_labour_force)),
    Completed_year_12_perc = as.numeric(as.character(Completed_year_12_perc)),
    Bachelor_degree_perc = as.numeric(as.character(Bachelor_degree_perc)),
    Graduate_diploma_perc = as.numeric(as.character(Graduate_diploma_perc)),
    Number_employee_jobs_professional_scientific_technical = as.numeric(as.character(Number_employee_jobs_professional_scientific_technical)),
    Persons_who_made_HELP_repayment = as.numeric(as.character(Persons_who_made_HELP_repayment)),
    National_parks = as.numeric(as.character(National_parks)),
    Nature_reserves = as.numeric(as.character(Nature_reserves)),
    Total_protected_land_area = as.numeric(as.character(Total_protected_land_area)),
    Persons_who_made_gift_or_donation = as.numeric(as.character(Persons_who_made_gift_or_donation)),
  )

#---- incorporating breed behavioral characteristic related data----

new_dog_data <- read.csv('dogs_cleaned.csv')

# Ensure the format of new_dog_data$Breed.Name matches frequency_data$nb_breed_name_unique
new_dog_data$Breed.Name <- tolower(new_dog_data$Breed.Name)  # Convert to lowercase
new_dog_data$Breed.Name <- trimws(new_dog_data$Breed.Name)   # Remove leading/trailing whitespace

frequency_data$nb_breed_name_unique <- tolower(frequency_data$nb_breed_name_unique)  # Convert to lowercase
frequency_data$nb_breed_name_unique <- trimws(frequency_data$nb_breed_name_unique)   # Remove leading/trailing whitespace

# Get unique breed names from both datasets
unique_frequency_breeds <- unique(frequency_data$nb_breed_name_unique)
unique_new_dog_breeds <- unique(new_dog_data$Breed.Name)

# Find breeds in frequency_data that are not in new_dog_data
missing_breeds <- setdiff(unique_frequency_breeds, unique_new_dog_breeds)

# Calculate the string distance matrix between missing breeds and new dog breeds
distance_matrix <- stringdistmatrix(missing_breeds, unique_new_dog_breeds, method = "jw")  # Jaro-Winkler method

# Find the index of the closest match for each missing breed
closest_match_indices <- apply(distance_matrix, 1, which.min)

# Get the closest matches from new_dog_data
closest_matches <- unique_new_dog_breeds[closest_match_indices]

# Create a data frame with missing breeds and their closest match
matching_df <- data.frame(
  Missing_Breeds = missing_breeds,
  Suggested_Matches = closest_matches,
  stringsAsFactors = FALSE
)

# Create a named vector for easier replacement
replacement_vector <- setNames(matching_df$Suggested_Matches, matching_df$Missing_Breeds)

# Replace values in 'nb_breed_name_unique' with the corresponding 'Suggested_Matches' values
frequency_data$nb_breed_name_unique <- ifelse(
  frequency_data$nb_breed_name_unique %in% matching_df$Missing_Breeds, 
  replacement_vector[frequency_data$nb_breed_name_unique], 
  frequency_data$nb_breed_name_unique
)

# Merge frequency_data with new_dog_data on breed names
frequency_data <- merge(frequency_data, new_dog_data, by.x = "nb_breed_name_unique", by.y = "Breed.Name", all.x = TRUE)

# Convert specified columns to factors
frequency_data$Dog.Breed.Group <- as.factor(frequency_data$Dog.Breed.Group)
frequency_data$Dog.Size <- as.factor(frequency_data$Dog.Size)
frequency_data$Height <- as.factor(frequency_data$Height)
frequency_data$Weight <- as.factor(frequency_data$Weight)
frequency_data$Life.Span <- as.factor(frequency_data$Life.Span)

# Assuming both 'claims' and 'frequency_data' datasets have a common 'exposure_id' column
severity_data <- claims %>%
  left_join(frequency_data, by = "exposure_id")

# Removing 'National_parks' and 'Nature_reserves' from frequency_data and severity_data
frequency_data <- frequency_data %>%
  select(-National_parks, -Nature_reserves)

severity_data <- severity_data %>%
  select(-National_parks, -Nature_reserves)

# Remove the specified columns from frequency_data and severity_data
columns_to_remove <- c('Total_protected_land_area', 'Persons_who_made_gift_or_donation', 
                       'Median_price_established_house_transfers', 'Value_private_sector_houses')

# Removing from frequency_data
frequency_data <- frequency_data %>%
  select(-all_of(columns_to_remove))

# Removing from severity_data
severity_data <- severity_data %>%
  select(-all_of(columns_to_remove))

# Remove rows where earned_units is 0 in frequency_data
frequency_data <- frequency_data[frequency_data$earned_units > 0, ]

```

After this, feature engineering was performed to create new interaction variables between breed data and policyholder information.

```{r}

# Define a function to add the exp_future_life column to a dataset
add_exp_future_life <- function(data) {
  if("Avg..Life.Span..years" %in% colnames(data) & "pet_age_months" %in% colnames(data)) {
    data$exp_future_life <- data$Avg..Life.Span..years - (data$pet_age_months / 12)
  } else {
    print("One or both columns 'Avg..Life.Span..years' and 'pet_age_months' are missing in the dataset.")
  }
  return(data)
}

# Apply the function to each dataset
frequency_data <- add_exp_future_life(frequency_data)
severity_data <- add_exp_future_life(severity_data)

# Define a function to add the new feature based on nb_address_type_adj
add_apartment_feature <- function(data) {
  data$happy_in_apartment <- ifelse(data$nb_address_type_adj == "Apartment", 
                             data$Adapts.Well.To.Apartment.Living, 
                             0)
  return(data)
}

# Apply the function to both frequency_data and frequency_data2
frequency_data <- add_apartment_feature(frequency_data)
severity_data <- add_apartment_feature(severity_data)

# Define a function to create the dog_safety interaction feature
add_dog_safety_feature <- function(data) {
  data$dog_safety <- ifelse(data$is_multi_pet_plan == TRUE, data$Dog.Friendly, 0)
  return(data)
}

# Apply the function to both frequency_data and severity_data
frequency_data <- add_dog_safety_feature(frequency_data)
severity_data <- add_dog_safety_feature(severity_data)

# Apply the function to both frequency_data and severity_data
frequency_data <- add_dog_safety_feature(frequency_data)
severity_data <- add_dog_safety_feature(severity_data)

# Define a function to add the expected_aggression feature
add_expected_aggression <- function(data) {
  data$expected_aggression <- ifelse(data$pet_de_sexed == TRUE, 
                                     0, 
                                     5 - data$All.Around.Friendliness)
  return(data)
}

# Apply the function to both frequency_data and severity_data
frequency_data <- add_expected_aggression(frequency_data)
severity_data <- add_expected_aggression(severity_data)

```

### Clustering

With the data prepared, clustering was performed to group condition categories; this formed the basis of our model approach.

```{r}

#----Clustering----

# Step 1: Count the number of claims for each exposure_id and condition_category
claim_counts2 <- claims %>%
  group_by(exposure_id, condition_category) %>%
  summarise(claim_count = n()) %>%
  ungroup()

# Merge with earned data to get earned_units
frequency_data_clus <- claim_counts2 %>%
  left_join(earned, by = "exposure_id") %>%
  select(exposure_id, condition_category, claim_count, earned_units)

# Calculate claims per earned_units for frequency
frequency_data_clus <- frequency_data_clus %>%
  mutate(frequency = claim_count / earned_units)

# Calculate mean and standard deviation of frequency for each condition_category
frequency_summary <- frequency_data_clus %>%
  group_by(condition_category) %>%
  summarise(
    mean_frequency = mean(frequency, na.rm = TRUE),
    sd_frequency = sd(frequency, na.rm = TRUE)
  )

# Calculate mean and standard deviation of claim_paid for severity_data
severity_summary <- severity_data %>%
  group_by(condition_category) %>%
  summarise(
    mean_severity = mean(claim_paid, na.rm = TRUE),
    sd_severity = sd(claim_paid, na.rm = TRUE)
  )

# Merging both summaries into a single data frame
condition_summary <- merge(frequency_summary, severity_summary, by = "condition_category")

# Step 1: Prepare the data with mean_frequency, mean_severity, and sd_severity
condition_data_scaled_3d <- scale(condition_summary[, c("mean_frequency", "mean_severity", "sd_severity")])

# Step 2: Perform K-means clustering with 3 centers (as determined earlier)
set.seed(123)
kmeans_3d <- kmeans(condition_data_scaled_3d, centers = 5, nstart = 25)

# Step 3: Calculate silhouette score for this clustering
silhouette_3d <- silhouette(kmeans_3d$cluster, dist(condition_data_scaled_3d))

# Step 4: Visualize the silhouette plot for this clustering
fviz_silhouette(silhouette_3d) + ggtitle("Silhouette Plot - Clustering with Mean Frequency, Mean Severity, and SD Severity")

# Step 5: Calculate the average silhouette width
avg_silhouette_3d <- mean(silhouette_3d[, 3])
cat("Average Silhouette Width (3D Clustering - Mean Frequency, Mean Severity, and SD Severity):", avg_silhouette_3d, "\n")

# Step 6: Add the cluster results to the original data
condition_summary$cluster <- as.factor(kmeans_3d$cluster)

# Step 7: Create the 3D scatter plot using the standardized values
plot_ly(
  data = condition_summary,
  x = ~mean_frequency,  # This is now the standardized mean_frequency
  y = ~mean_severity,   # Standardized mean_severity
  z = ~sd_severity,     # Standardized sd_severity
  color = ~cluster,     # Cluster will be represented by color
  size = ~mean_severity,  # Size represents the mean_severity
  text = ~condition_category,  # Condition category as hover text
  type = 'scatter3d',   # 3D scatter plot
  mode = 'markers',
  marker = list(sizemode = 'diameter')  # Adjust size by diameter
) %>%
  layout(
    title = "3D Scatter Plot of Condition Categories",
    scene = list(
      xaxis = list(title = 'Mean Frequency'),
      yaxis = list(title = 'Mean Severity'),
      zaxis = list(title = 'SD Severity')
    )
  )



# Step 1: Count the number of claims for each exposure_id and condition_category
claim_counts2 <- claims %>%
  group_by(exposure_id, condition_category) %>%
  summarise(claim_count = n()) %>%
  ungroup()

frequency_data_clus <- claim_counts2 %>%
  left_join(earned, by = "exposure_id") %>%
  select(exposure_id, condition_category, claim_count)

# Calculate mean and standard deviation of claim_count for each condition_category
frequency_summary <- frequency_data_clus %>%
  group_by(condition_category) %>%
  summarise(
    mean_frequency = mean(claim_count, na.rm = TRUE),
    sd_frequency = sd(claim_count, na.rm = TRUE)
  )

# Calculate mean and standard deviation of claim_paid for severity_data
severity_summary <- severity_data %>%
  group_by(condition_category) %>%
  summarise(
    mean_severity = mean(claim_paid, na.rm = TRUE),
    sd_severity = sd(claim_paid, na.rm = TRUE)
  )

# Merging both summaries into a single data frame
condition_summary <- merge(frequency_summary, severity_summary, by = "condition_category")

# Step 1: Prepare the data with mean_frequency, mean_severity, and sd_severity
condition_data_scaled_3d <- scale(condition_summary[, c("mean_frequency", "mean_severity", "sd_severity")])

# Step 2: Perform K-means clustering with 3 centers (as determined earlier)
set.seed(123)
kmeans_3d <- kmeans(condition_data_scaled_3d, centers = 5, nstart = 25)

# Step 3: Calculate silhouette score for this clustering
silhouette_3d <- silhouette(kmeans_3d$cluster, dist(condition_data_scaled_3d))

# Step 4: Visualize the silhouette plot for this clustering
fviz_silhouette(silhouette_3d) + ggtitle("Silhouette Plot - Clustering with Mean Frequency, Mean Severity, and SD Severity")

# Step 5: Calculate the average silhouette width
avg_silhouette_3d <- mean(silhouette_3d[, 3])
cat("Average Silhouette Width (3D Clustering - Mean Frequency, Mean Severity, and SD Severity):", avg_silhouette_3d, "\n")

# Step 6: Add the cluster results to the original data
condition_summary$cluster <- as.factor(kmeans_3d$cluster)

# Step 7: Create the 3D scatter plot using the standardized values
plot_ly(
  data = condition_summary,
  x = ~mean_frequency,  # This is now the standardized mean_frequency
  y = ~mean_severity,   # Standardized mean_severity
  z = ~sd_severity,     # Standardized sd_severity
  color = ~cluster,     # Cluster will be represented by color
  size = ~mean_severity,  # Size represents the mean_severity
  text = ~condition_category,  # Condition category as hover text
  type = 'scatter3d',   # 3D scatter plot
  mode = 'markers',
  marker = list(sizemode = 'diameter')  # Adjust size by diameter
) %>%
  layout(
    title = "3D Scatter Plot of Condition Categories",
    scene = list(
      xaxis = list(title = 'Mean Frequency'),
      yaxis = list(title = 'Mean Severity'),
      zaxis = list(title = 'SD Severity')
    )
  )





# Step 1: Define a range of k values
k_values <- 1:10

# Step 2: Initialize a vector to hold the total WSS for each k
wss_values <- numeric(length(k_values))

# Step 3: Calculate WSS for each k
for (k in k_values) {
  kmeans_model <- kmeans(condition_data_scaled_3d, centers = k, nstart = 25)
  wss_values[k] <- kmeans_model$tot.withinss
}

# Create the elbow plot using ggplot
elbow_plot <- ggplot(data = data.frame(k = k_values, wss = wss_values), aes(x = k, y = wss)) +
  geom_point(color = "skyblue", size = 3) + 
  geom_line(color = "darkblue") + 
  labs(x = "Number of Clusters (k)", y = "Total Within-Cluster Sum of Squares (WSS)") +
  scale_x_continuous(breaks = 1:10) +  # Specify breaks for x-axis from 1 to 10
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5), 
        legend.position = "none")  # Set angle to 0 for horizontal labels and remove legend

# Save the elbow plot to a file
ggsave(filename = "elbow_plot.png", height = 11, width = 12, units = 'cm', dpi = 1080, plot = elbow_plot)


# Step 1: Combine clusters 1, 2, and 3 into a single cluster (renamed as "1")
condition_summary$merged_cluster <- ifelse(condition_summary$cluster %in% c(1, 2, 3), 1, condition_summary$cluster)

# Step 2: Convert the new merged cluster column to a factor
condition_summary$merged_cluster <- as.factor(condition_summary$merged_cluster)

# Step 3: Create a new 3D scatter plot with the merged clusters
plot_ly(
  data = condition_summary,
  x = ~mean_frequency,  # Standardized mean_frequency
  y = ~mean_severity,   # Standardized mean_severity
  z = ~sd_severity,     # Standardized sd_severity
  color = ~merged_cluster,  # Use the new merged cluster
  size = ~mean_severity,    # Size represents the mean_severity
  text = ~condition_category,  # Hover text with condition_category
  type = 'scatter3d',         # 3D scatter plot
  mode = 'markers',
  marker = list(sizemode = 'diameter')  # Adjust size by diameter
) %>%
  layout(
    title = "3D Scatter Plot of Merged Clusters (1, 2, 3 grouped together)",
    scene = list(
      xaxis = list(title = 'Mean Frequency'),
      yaxis = list(title = 'Mean Severity'),
      zaxis = list(title = 'SD Severity')
    )
  )

# Load necessary libraries
library(cluster)  # For silhouette calculation
library(dplyr)    # For data manipulation
library(entropy)  # For entropy calculation

# Step 1: Calculate the silhouette score
# Ensure the merged cluster is in the right format
condition_summary$merged_cluster <- as.numeric(condition_summary$merged_cluster)

# Calculate silhouette score
silhouette_values <- silhouette(condition_summary$merged_cluster, dist(condition_data_scaled_3d))
avg_silhouette_score <- mean(silhouette_values[, 3])  # Average silhouette width

cat("Average Silhouette Score (after merging clusters):", avg_silhouette_score, "\n")

# Step 2: Calculate entropy
# Create a contingency table
contingency_table <- table(condition_summary$merged_cluster, condition_summary$condition_category)

# Calculate probabilities for each cluster
probabilities <- prop.table(contingency_table, margin = 1)

# Calculate entropy for each cluster
entropy_values <- apply(probabilities, 1, function(p) {
  -sum(p * log2(p + (p == 0)))  # Avoid log(0) by adding a small value
})

# Calculate the overall entropy
overall_entropy <- sum(entropy_values * rowSums(contingency_table) / sum(contingency_table))

cat("Overall Entropy (after merging clusters):", overall_entropy, "\n")



#----Assign clusters----

claim_counts_with_cluster <- claim_counts2 %>%
  left_join(condition_summary %>% 
              mutate(merged_cluster = ifelse(cluster %in% c(1, 2, 3), 1, cluster)) %>%
              select(condition_category, merged_cluster), 
            by = "condition_category") %>%
  group_by(exposure_id, merged_cluster) %>%
  summarise(claim_count = sum(claim_count, na.rm = TRUE)) %>%
  ungroup()

# Step 2: Pivot the data to create columns for each merged cluster
cluster_frequencies <- claim_counts_with_cluster %>%
  spread(key = merged_cluster, value = claim_count, fill = 0)

# Rename the cluster columns to cat1 (for merged 1, 2, 3), cat2, and cat3 (for original clusters 4 and 5)
names(cluster_frequencies)[2:4] <- c("cat1", "cat2", "cat3")

# Step 3: Join the cluster frequencies back to the 'frequency_data' dataset
frequency_data <- frequency_data %>%
  left_join(cluster_frequencies, by = "exposure_id")

# Step 4: Handle cases where claim_count = 0 by setting cat1, cat2, and cat3 to 0
frequency_data <- frequency_data %>%
  mutate(
    cat1 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat1),
    cat2 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat2),
    cat3 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat3)
  )

frequency_data <- frequency_data %>%
  mutate(
    claim_rate_cat1 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat1 / earned_units),
    claim_rate_cat2 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat2 / earned_units),
    claim_rate_cat3 = ifelse(is.na(claim_count) | claim_count == 0, 0, cat3 / earned_units)
  )

```

With this completed, further feature engineering was performed.

```{r}

#----Feature engineering on pre-provided features----

frequency_data$nb_policy_first_inception_date <- as.Date(frequency_data$nb_policy_first_inception_date, format="%Y-%m-%d")

# Extract year and month from the inception date
frequency_data <- frequency_data %>%
  mutate(inception_year = as.numeric(format(nb_policy_first_inception_date, "%Y")),
         inception_month = as.numeric(format(nb_policy_first_inception_date, "%m")))

# Find the maximum inception date in the dataset
max_inception_date <- max(frequency_data$nb_policy_first_inception_date, na.rm = TRUE)

# Calculate the number of months between the inception date and the maximum date in the variable
frequency_data <- frequency_data %>%
  mutate(policy_age_months = as.numeric(difftime(max_inception_date, nb_policy_first_inception_date, units = "days")) / 30)

# Create a season feature based on the month of inception
frequency_data <- frequency_data %>%
  mutate(season_inception = case_when(
    inception_month %in% c(12, 1, 2) ~ "Summer",
    inception_month %in% c(3, 4, 5) ~ "Autumn",
    inception_month %in% c(6, 7, 8) ~ "Winter",
    inception_month %in% c(9, 10, 11) ~ "Spring"
  ))

# Create policy age groups relative to the maximum inception date
frequency_data <- frequency_data %>%
  mutate(policy_age_group = case_when(
    policy_age_months <= 6 ~ "Very New",
    policy_age_months <= 12 ~ "New",
    policy_age_months <= 36 ~ "Mid",
    TRUE ~ "Established"
  ))

# Create an indicator for recently incepted policies (e.g., within the last 6 months relative to the maximum inception date)
frequency_data <- frequency_data %>%
  mutate(is_recent_policy = ifelse(policy_age_months <= 6, 1, 0))

# Calculate days since the start of the inception year
frequency_data <- frequency_data %>%
  mutate(days_since_start_of_year = as.numeric(difftime(nb_policy_first_inception_date, as.Date(paste0(inception_year, "-01-01")), units = "days")))

# Calculate the number of months since the start of the coverage
frequency_data <- frequency_data %>%
  mutate(months_since_UW = as.numeric(difftime(Sys.Date(), UW_Date, units = "days")) / 30)

# Convert to Date if needed
frequency_data$lead_date_day <- as.Date(frequency_data$lead_date_day, format="%Y-%m-%d")
frequency_data$nb_policy_first_inception_date <- as.Date(frequency_data$nb_policy_first_inception_date, format="%Y-%m-%d")

# Calculate the time difference between lead_date_day and nb_policy_first_inception_date in days
frequency_data <- frequency_data %>%
  mutate(time_to_inception_days = as.numeric(difftime(nb_policy_first_inception_date, lead_date_day, units = "days")))

### Note tenure is captured by earned_units

cor(frequency_data$tenure, frequency_data$earned_units)

frequency_data2 <- frequency_data %>%
  select(exp_future_life, exposure_id, inception_year, policy_age_months, season_inception, earned_units,
         policy_age_group, 
         claim_count, cat1, cat2, cat3, claim_rate_cat1, claim_rate_cat2, claim_rate_cat3,
         nb_contribution_excess, nb_contribution, nb_excess,
         pet_age_months, nb_average_breed_size,
         is_multi_pet_plan, pet_de_sexed, pet_gender,
         nb_address_type_adj, nb_breed_type, owner_age_years,
         Total_number_bus, Agriculture_forestry_fishing_per,
         Health_care_social_assistance_per, Completed_year_12_perc,
         Average_household_size, Participation_rate,
         Median_employee_income_SA3, Dog.Breed.Group,
         Sensitivity.Level,happy_in_apartment, dog_safety,
         Tolerates.Cold.Weather, Tolerates.Hot.Weather,
         Kid.Friendly, Friendly.Toward.Strangers,
         Drooling.Potential, Easy.To.Groom, General.Health,
         Potential.For.Weight.Gain, Intelligence,
         Potential.For.Mouthiness, Prey.Drive, expected_aggression,
         Tendency.To.Bark.Or.Howl, Wanderlust.Potential,
         Energy.Level, Intensity)

# Assuming frequency_data is your original data frame
frequency_data <- frequency_data %>%
  select(exp_future_life, inception_year, policy_age_months, season_inception, earned_units,
         policy_age_group, 
         claim_count, cat1, cat2, cat3, claim_rate_cat1, claim_rate_cat2, claim_rate_cat3,
         nb_contribution_excess, nb_contribution, nb_excess,
         pet_age_months, nb_average_breed_size,
         is_multi_pet_plan, pet_de_sexed, pet_gender,
         nb_address_type_adj, nb_breed_type, owner_age_years,
         Total_number_bus, Agriculture_forestry_fishing_per,
         Health_care_social_assistance_per, Completed_year_12_perc,
         Average_household_size, Participation_rate,
         Median_employee_income_SA3, Dog.Breed.Group,
         Sensitivity.Level,dog_safety, happy_in_apartment,
         Tolerates.Cold.Weather, Tolerates.Hot.Weather,
         Kid.Friendly, Friendly.Toward.Strangers,
         Drooling.Potential, Easy.To.Groom, General.Health,
         Potential.For.Weight.Gain, Intelligence,
         Potential.For.Mouthiness, Prey.Drive, expected_aggression,
         Tendency.To.Bark.Or.Howl, Wanderlust.Potential,
         Energy.Level, Intensity)

```

With these additional factors, preliminary feature selection was performed, removing variables with high correlations, and those we deemed unnecessary.

```{r}

# Step 1: Select only numeric columns for the correlation matrix
numeric_data <- frequency_data[sapply(frequency_data, is.numeric)]

# Step 2: Calculate the correlation matrix
cor_matrix <- cor(numeric_data, use = "complete.obs")  # 'complete.obs' excludes NA values

# Step 3: Find correlations greater than 0.5 (ignoring self-correlations)
high_corr <- which(abs(cor_matrix) > 0.5 & abs(cor_matrix) < 1, arr.ind = TRUE)

# Display the pairs of variables with correlation greater than 0.5
high_corr_pairs <- data.frame(
  Variable1 = rownames(cor_matrix)[high_corr[, 1]],
  Variable2 = colnames(cor_matrix)[high_corr[, 2]],
  Correlation = cor_matrix[high_corr]
)

# Print the pairs of highly correlated variables
print(high_corr_pairs)

# Remove the 'policy_age_months' and 'inception_year' columns from frequency_data
frequency_data <- frequency_data[, !(names(frequency_data) %in% c("policy_age_months", "inception_year"))]

# Remove the 'Dog.Breed.Group', 'season_inception', and 'risk_group' columns from frequency_data
frequency_data <- frequency_data[, !(names(frequency_data) %in% c("Dog.Breed.Group", "season_inception"))]
frequency_data2 <- frequency_data2[, !(names(frequency_data2) %in% c("Dog.Breed.Group", "season_inception"))]

```

### Data Splitting

Finally, data was split into test and training sets, upon which the models could be trained and evaluated.

```{r}

#----Creating data splits----

# Load required libraries
library(dplyr)

# List of exceptions (variables not to convert)
exceptions <- c("Avg..Life.Span..years")

# Identify the columns with dots (.) in their names that are not exceptions
cols_to_change <- colnames(frequency_data) %>%
  grep("\\.", ., value = TRUE) %>%
  setdiff(exceptions)

# Add "Intensity" and "Intelligence" to the list of variables to be converted
cols_to_change <- union(cols_to_change, c("Intensity", "Intelligence"))

# Convert these columns to ordinal variables with 5 levels
frequency_data <- frequency_data %>%
  mutate(across(all_of(cols_to_change), ~ cut(., breaks = 5, labels = 1:5, ordered_result = TRUE)))

# Display the structure of the modified dataset
str(frequency_data)


# Convert these columns to ordinal variables with 5 levels
severity_data <- severity_data %>%
  mutate(across(all_of(cols_to_change), ~ cut(., breaks = 5, labels = 1:5, ordered_result = TRUE)))

# Display the structure of the modified dataset
str(severity_data)

# Convert these columns to ordinal variables with 5 levels
frequency_data2 <- frequency_data2 %>%
  mutate(across(all_of(cols_to_change), ~ cut(., breaks = 5, labels = 1:5, ordered_result = TRUE)))

# Display the structure of the modified dataset
str(frequency_data2)

# Remove rows where Health_care_social_assistance_per or Median_employee_income_SA3 is NA - avoids issues later on when modelling
frequency_data <- frequency_data %>%
  filter(!is.na(Health_care_social_assistance_per) & !is.na(Median_employee_income_SA3) & !is.na(cat1) & !is.na(Drooling.Potential))

frequency_data2 <- frequency_data2 %>%
  filter(!is.na(Health_care_social_assistance_per) & !is.na(Median_employee_income_SA3) & !is.na(cat1) & !is.na(Drooling.Potential))

# Convert character columns to factors
frequency_data$policy_age_group <- as.factor(frequency_data$policy_age_group)
frequency_data2$policy_age_group <- as.factor(frequency_data2$policy_age_group)

# Set a seed for reproducibility
set.seed(123)

# Create a single train-test split for the frequency data (80% train, 20% test)
train_index <- sample(seq_len(nrow(frequency_data)), size = 0.8 * nrow(frequency_data))
train_data <- frequency_data[train_index, ]
test_data <- frequency_data[-train_index, ]

# Extract year and month from the inception date
severity_data <- severity_data %>%
  mutate(inception_month = as.numeric(format(nb_policy_first_inception_date, "%m")))

# Calculate the number of months between the inception date and the maximum date in the variable
severity_data <- severity_data %>%
  mutate(policy_age_months = as.numeric(difftime(max_inception_date, nb_policy_first_inception_date, units = "days")) / 30)

# Create policy age groups relative to the maximum inception date
severity_data <- severity_data %>%
  mutate(policy_age_group = case_when(
    policy_age_months <= 6 ~ "Very New",
    policy_age_months <= 12 ~ "New",
    policy_age_months <= 36 ~ "Mid",
    TRUE ~ "Established"
  ))

# Extract year and month of coverage
severity_data <- severity_data %>%
  mutate(UW_month = as.numeric(format(UW_Date, "%m")))

# Join severity_data with the merged cluster assignment from condition_summary by condition_category
severity_data <- severity_data %>%
  left_join(condition_summary %>% 
              mutate(merged_cluster = ifelse(cluster %in% c(1, 2, 3), 1, cluster)) %>%
              select(condition_category, merged_cluster), 
            by = "condition_category")

severity_data <- severity_data %>%
  select(exp_future_life, earned_units,happy_in_apartment, dog_safety,
         policy_age_group, merged_cluster, 
         claim_paid, pet_age_months, nb_average_breed_size,
    nb_contribution_excess, nb_contribution, nb_excess,
    is_multi_pet_plan, pet_de_sexed, pet_gender,
    nb_address_type_adj, nb_breed_type, owner_age_years,
    Total_number_bus, Agriculture_forestry_fishing_per,
    Health_care_social_assistance_per, Completed_year_12_perc,
    Average_household_size, Participation_rate,
    Median_employee_income_SA3,
    Sensitivity.Level,expected_aggression,
    Tolerates.Cold.Weather, Tolerates.Hot.Weather,
    Kid.Friendly, Friendly.Toward.Strangers,
    Drooling.Potential, Easy.To.Groom, General.Health,
    Potential.For.Weight.Gain, Intelligence,
    Potential.For.Mouthiness, Prey.Drive,
    Tendency.To.Bark.Or.Howl, Wanderlust.Potential,
    Energy.Level, Intensity
  )

# Step 1: Remove rows where Energy.Level or Health_care_social_assistance_per are NA
severity_data <- severity_data %>%
  filter(!is.na(Energy.Level) & !is.na(Health_care_social_assistance_per))

# Convert 'policy_age_group' and 'UW_season' to factors in severity_data
severity_data$policy_age_group <- as.factor(severity_data$policy_age_group)

# Subset the severity data based on the merged clusters
severity_data_cat1 <- severity_data %>% filter(merged_cluster == 1)
severity_data_cat2 <- severity_data %>% filter(merged_cluster == 4)
severity_data_cat3 <- severity_data %>% filter(merged_cluster == 5)

# Apply the same train-test split to the severity data for Cat1
train_data_cat1 <- severity_data_cat1[rownames(severity_data_cat1) %in% train_index, ]
test_data_cat1 <- severity_data_cat1[!rownames(severity_data_cat1) %in% train_index, ]

# Apply the same train-test split to the severity data for Cat2
train_data_cat2 <- severity_data_cat2[rownames(severity_data_cat2) %in% train_index, ]
test_data_cat2 <- severity_data_cat2[!rownames(severity_data_cat2) %in% train_index, ]

# Apply the same train-test split to the severity data for Cat3
train_data_cat3 <- severity_data_cat3[rownames(severity_data_cat3) %in% train_index, ]
test_data_cat3 <- severity_data_cat3[!rownames(severity_data_cat3) %in% train_index, ]

# Remove the 'merged_cluster' and 'earned_units' columns from the severity training and test datasets

# Cat1
train_data_cat1 <- train_data_cat1 %>% select(-merged_cluster, -earned_units)
test_data_cat1 <- test_data_cat1 %>% select(-merged_cluster, -earned_units)

# Cat2
train_data_cat2 <- train_data_cat2 %>% select(-merged_cluster, -earned_units)
test_data_cat2 <- test_data_cat2 %>% select(-merged_cluster, -earned_units)

# Cat3
train_data_cat3 <- train_data_cat3 %>% select(-merged_cluster, -earned_units)
test_data_cat3 <- test_data_cat3 %>% select(-merged_cluster, -earned_units)

# Check if all exposures in test_data are covered by test_data_cat1, test_data_cat2, and test_data_cat3
all_exposures_test_data <- test_data$exposure_id
cat1_exposures <- test_data_cat1$exposure_id
cat2_exposures <- test_data_cat2$exposure_id
cat3_exposures <- test_data_cat3$exposure_id

# Check if the combined exposures of cat1, cat2, and cat3 are the same as test_data
setequal(all_exposures_test_data, union(union(cat1_exposures, cat2_exposures), cat3_exposures))


# Set a seed for reproducibility
set.seed(123)

# Create a single train-test split for the frequency data (80% train, 20% test)
train_index <- sample(seq_len(nrow(frequency_data)), size = 0.8 * nrow(frequency_data))

test_data_prem <- frequency_data2[-train_index, ]

test_data_prem$exp_future_life
```

### Modelling

First, train EBM models on the frequency data.

```{r}
#----Frequency Models----

# Load required libraries
library(caret)
library(reticulate)

# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to build EBM, calculate RMSE, residuals, and print top 5/bottom 10 features
evaluate_ebm_rmse <- function(train_data, test_data, target_var, vars_to_remove) {
  # Remove specified variables from training and test sets
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  test_data <- test_data[, !names(test_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training and testing data
  train_data <- na.omit(train_data)
  test_data <- na.omit(test_data)
  
  # Prepare predictor variables and target variables for training and testing
  X_train <- train_data[, !names(train_data) %in% target_var]  # Predictor variables for training
  y_train <- train_data[[target_var]]  # Target variable for training
  
  X_test <- test_data[, !names(test_data) %in% target_var]  # Predictor variables for testing
  y_test <- test_data[[target_var]]  # Target variable for testing
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  # Make predictions using the test data
  predictions <- ebm$predict(X_test)
  
  # Calculate residuals
  residuals <- y_test - predictions
  
  # Calculate RMSE using the test data
  rmse <- sqrt(mean(residuals^2))
  
  # Get feature importance scores
  global_explanation <- ebm$explain_global()
  feature_importance <- global_explanation$data()$scores
  feature_names <- global_explanation$data()$names
  
  # Convert feature importance to numeric
  feature_importance <- as.numeric(feature_importance)
  
  # Combine feature names and importance into a data frame
  importance_df <- data.frame(Feature = feature_names, Importance = feature_importance)
  
  # Sort by importance (descending order)
  importance_df <- importance_df[order(-importance_df$Importance), ]
  
  # Print top 5 important features
  cat("Top 5 Important Features:\n")
  print(importance_df[ , ])
  
  # Print bottom 10 important features
  cat("\nBottom 10 Important Features:\n")
  print(importance_df[(nrow(importance_df)-9):nrow(importance_df), ])
  
  return(list(rmse = rmse, residuals = residuals))
}

# Build EBM and evaluate RMSE for different target variables
result_cat1 <- evaluate_ebm_rmse(train_data, test_data, "claim_rate_cat1", c("claim_count", "earned_units", "cat1", "cat2", "cat3", 
                                                                             "claim_rate_cat2", "claim_rate_cat3", "Drooling.Potential",
                                                                             "nb_contribution", "Energy.Level", "policy_age_group", "is_multi_pet_plan",
                                                                             "pet_de_sexed"))

cat("RMSE for Cat1 prediction:", result_cat1$rmse, "\n")


## Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict
evaluate_and_predict_ebm <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Prepare predictor variables and target variables for training
  X_train <- train_data[, !names(train_data) %in% target_var]  # Predictor variables for training
  y_train <- train_data[[target_var]]  # Target variable for training
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, fill with NA (or 0 if appropriate)
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA  # Or 0 depending on the nature of the variable
  }
  
  # Ensure test_data_prem_adjusted has no NAs
  test_data_prem_adjusted <- na.omit(test_data_prem_adjusted)
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var]  # Predictor variables for testing
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_prem)
  
  # Replace negative predictions with the smallest positive value from the predictions
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}

# Define the list of variables to remove (as used before)
vars_to_remove <- c("claim_count", "earned_units", "cat1", "cat2", "cat3", 
                    "claim_rate_cat2", "claim_rate_cat3", "Drooling.Potential",
                    "nb_contribution", "Energy.Level", "policy_age_group", 
                    "is_multi_pet_plan", "pet_de_sexed")

# Call the function to predict for test_data_prem using the EBM model for Cat1
predicted_frequencies_prem_cat1 <- evaluate_and_predict_ebm(train_data = train_data, 
                                                            test_data_prem = test_data_prem, 
                                                            target_var = "claim_rate_cat1", 
                                                            vars_to_remove = vars_to_remove)


result_cat2 <- evaluate_ebm_rmse(train_data, test_data, "claim_rate_cat2", c("claim_count", "earned_units", "cat1", "cat2", "cat3", "claim_rate_cat1", "claim_rate_cat3", 
                                                                             "is_multi_pet_plan", "pet_de_sexed", "UW_season", "Energy.Level", "policy_age_group"))
cat("RMSE for Cat2 prediction:", result_cat2$rmse, "\n")


# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict
evaluate_and_predict_ebm_cat2 <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Prepare predictor variables and target variables for training
  X_train <- train_data[, !names(train_data) %in% target_var]  # Predictor variables for training
  y_train <- train_data[[target_var]]  # Target variable for training
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, fill with NA (or 0 if appropriate)
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA  # Or 0 depending on the nature of the variable
  }
  
  # Ensure test_data_prem_adjusted has no NAs
  test_data_prem_adjusted <- na.omit(test_data_prem_adjusted)
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var]  # Predictor variables for testing
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_prem)
  
  # Replace negative predictions with the smallest positive value from the predictions
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}

# Define the list of variables to remove for Cat2 model
vars_to_remove_cat2 <- c("claim_count", "earned_units", "cat1", "cat2", "cat3", 
                         "claim_rate_cat1", "claim_rate_cat3", "is_multi_pet_plan", 
                         "pet_de_sexed", "UW_season", "Energy.Level", "policy_age_group")

# Call the function to predict for test_data_prem using the EBM model for Cat2
predicted_frequencies_prem_cat2 <- evaluate_and_predict_ebm_cat2(train_data = train_data, 
                                                                 test_data_prem = test_data_prem, 
                                                                 target_var = "claim_rate_cat2", 
                                                                 vars_to_remove = vars_to_remove_cat2)


result_cat3 <- evaluate_ebm_rmse(train_data, test_data, "claim_rate_cat3", c("cat1", "cat2", "cat3", "earned_units", "claim_count", "claim_rate_cat2", "claim_rate_cat1",
                                                                             "UW_season", "is_multi_pet_plan", "policy_age_group", "Energy.Level", "Intelligence", "Kid.Friendly",
                                                                             "pet_gender", "Sensitivity.Level"))
cat("RMSE for Cat3 prediction:", result_cat3$rmse, "\n")



# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict
evaluate_and_predict_ebm_cat3 <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Prepare predictor variables and target variables for training
  X_train <- train_data[, !names(train_data) %in% target_var]  # Predictor variables for training
  y_train <- train_data[[target_var]]  # Target variable for training
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, fill with NA (or 0 if appropriate)
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA  # Or 0 depending on the nature of the variable
  }
  
  # Ensure test_data_prem_adjusted has no NAs
  test_data_prem_adjusted <- na.omit(test_data_prem_adjusted)
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var]  # Predictor variables for testing
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_prem)
  
  # Replace negative predictions with the smallest positive value from the predictions
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}

# Define the list of variables to remove for Cat3 model
vars_to_remove_cat3 <- c("cat1", "cat2", "cat3", "earned_units", "claim_count", "claim_rate_cat2", 
                         "claim_rate_cat1", "UW_season", "is_multi_pet_plan", "policy_age_group", 
                         "Energy.Level", "Intelligence", "Kid.Friendly", "pet_gender", "Sensitivity.Level")

# Call the function to predict for test_data_prem using the EBM model for Cat3
predicted_frequencies_prem_cat3 <- evaluate_and_predict_ebm_cat3(train_data = train_data, 
                                                                 test_data_prem = test_data_prem, 
                                                                 target_var = "claim_rate_cat3", 
                                                                 vars_to_remove = vars_to_remove_cat3)
```

Next, train the group severity models.

```{r}
#----Severity models----

# Load necessary libraries
library(reticulate)

# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Variables to remove from the model
vars_to_remove <- c("Sensitivity.Level", "Kid.Friendly", "Drooling.Potential", "Potential.For.Mouthiness", "Avg..Life.Span..years", "Friendly.Toward.Strangers",
                    "Energy.Level", "Easy.To.Groom", "Dog.Friendly", "Intelligence", "Prey.Drive", "pet_gender", "Median_employee_income_SA3", "Intensity", "Tolerates.Hot.Weather",
                    "nb_address_type_adj", "inception_month", "owner_age_years", "Total_number_bus", "policy_age_group")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, test_data, vars_to_remove) {
  # Remove the specified variables from training and test datasets
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  test_data <- test_data[, !names(test_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variables for training and testing
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster")]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat1 <- train_ebm_model(train_data_cat1, test_data_cat1, vars_to_remove)

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat1, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem1 <- predict_claim_paid(ebm_model = ebm_model_cat1, 
                                                 test_data_prem = test_data_prem, 
                                                 train_data_cat1 = train_data_cat1, 
                                                 vars_to_remove = vars_to_remove)




# Variables to remove from the model
vars_to_remove <- c("UW_season", "Intelligence", "Kid.Friendly", "Average_household_size", "nb_address_type_adj", "pet_de_sexed",
                    "Sensitivity.Level", "Friendly.Toward.Strangers", "Tendency.To.Bark.Or.Howl", "Energy.Level", "Prey.Drive", "Easy.To.Groom",
                    "Potential.For.Weight.Gain", "Tolerates.Cold.Weather", "Total_number_bus")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, test_data, vars_to_remove) {
  # Remove the specified variables from training and test datasets
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  test_data <- test_data[, !names(test_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variables for training and testing
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster")]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat2 <- train_ebm_model(train_data_cat2, test_data_cat2, vars_to_remove)

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat2, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem2 <- predict_claim_paid(ebm_model = ebm_model_cat2, 
                                                 test_data_prem = test_data_prem, 
                                                 train_data_cat2 = train_data_cat2, 
                                                 vars_to_remove = vars_to_remove)



# Variables to remove from the model
vars_to_remove <- c("Potential.For.Weight.Gain", "Drooling.Potential", "Kid.Friendly", "owner_age_years", "Intelligence", "pet_gender",
                    "nb_address_type_adj", "UW_season", "Sensitivity.Level")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, test_data, vars_to_remove) {
  # Remove the specified variables from training and test datasets
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  test_data <- test_data[, !names(test_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variables for training and testing
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster")]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat3 <- train_ebm_model(train_data_cat3, test_data_cat3, vars_to_remove)


# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat3, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem3 <- predict_claim_paid(ebm_model = ebm_model_cat3, 
                                                 test_data_prem = test_data_prem, 
                                                 train_data_cat3 = train_data_cat3, 
                                                 vars_to_remove = vars_to_remove)

```

Then, train a model on overall severity (to be used in the credibility model).

```{r}

#----Total severity model----
# Variables to remove from the model
vars_to_remove <- c("UW_season")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, test_data, vars_to_remove) {
  # Remove the specified variables from training and test datasets
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  test_data <- test_data[, !names(test_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variables for training and testing
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster")]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_total <- train_ebm_model(train_data_total, test_data_total, vars_to_remove)

### Predict on test_data_prem

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat2, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_total <- predict_claim_paid(ebm_model = ebm_model_total, 
                                                 test_data_prem = test_data_prem, 
                                                 train_data_cat2 = train_data_total, 
                                                 vars_to_remove = vars_to_remove)

```

Now, construct the credibility model and calculate credibility factors for each of the condition groups.

```{r}

#----Credibility Adjustment----

## Bulmann-Straub Approach
# Function to calculate scaled credibility factors based on the number of rows in the training data
calculate_scaled_credibility <- function(train_data, overall_data, response_variable, total_rows) {
  # Number of observations in the category
  n <- nrow(train_data)
  
  # Mean of the response variable for individual model
  mean_individual <- mean(train_data[[response_variable]], na.rm = TRUE)
  
  # Overall mean of the response variable
  mean_overall <- mean(overall_data[[response_variable]], na.rm = TRUE)
  
  # Within-group variance (individual risk category)
  var_within <- var(train_data[[response_variable]], na.rm = TRUE)
  
  # Between-group variance (variance between individual mean and overall mean)
  var_between <- mean((mean_individual - mean_overall)^2)
  
  # Unscaled credibility factor Z (traditional formula)
  Z <- n / (n + (var_within / var_between))
  
  # Scale the credibility factor by the number of rows relative to the total
  Z_scaled <- Z
  
  return(list(credibility_factor = Z_scaled, mean_individual = mean_individual, mean_overall = mean_overall))
}

# Total number of rows in the overall data
total_rows <- nrow(train_data_total)

# Example calculation for train_data_cat1, train_data_cat2, train_data_cat3
credibility_cat1 <- calculate_scaled_credibility(train_data_cat1, train_data_total, "claim_paid", total_rows)
credibility_cat2 <- calculate_scaled_credibility(train_data_cat2, train_data_total, "claim_paid", total_rows)
credibility_cat3 <- calculate_scaled_credibility(train_data_cat3, train_data_total, "claim_paid", total_rows)

# Scaled Credibility factors for each category
Z_cat1 <- credibility_cat1$credibility_factor
Z_cat2 <- credibility_cat2$credibility_factor
Z_cat3 <- credibility_cat3$credibility_factor

```

Finally, use the models on the test data to calculate RMSE and Gini.

```{r}

# Adjusted severity calculation using the scaled credibility factors
adjusted_severity_cat1 <- Z_cat1 * predicted_claim_paid_prem1 + (1 - Z_cat1) * predicted_claim_paid_total
adjusted_severity_cat2 <- Z_cat2 * predicted_claim_paid_prem2 + (1 - Z_cat2) * predicted_claim_paid_total
adjusted_severity_cat3 <- Z_cat3 * predicted_claim_paid_prem3 + (1 - Z_cat3) * predicted_claim_paid_total


predicted_frequencies_prem_cat1
predicted_frequencies_prem_cat2
predicted_frequencies_prem_cat3

adjusted_severity_cat1
adjusted_severity_cat2
adjusted_severity_cat3

# Multiply corresponding entries
result_cat1 <- predicted_frequencies_prem_cat1 * adjusted_severity_cat1
result_cat2 <- predicted_frequencies_prem_cat2 * adjusted_severity_cat2
result_cat3 <- predicted_frequencies_prem_cat3 * adjusted_severity_cat3

# Perform the element-wise sum
total_result <- result_cat1 + result_cat2 + result_cat3

hist(total_result)

set.seed(123)
# Create a single train-test split for the frequency data (80% train, 20% test)
train_index <- sample(seq_len(nrow(frequency_data)), size = 0.8 * nrow(frequency_data))
train_data <- frequency_data[train_index, ]
test_data <- frequency_data[-train_index, ]

test_data_prem <- frequency_data2[-train_index, ]

# Step 1: Summarize claim_paid in the claims dataset
claims_summary <- claims %>%
  group_by(exposure_id) %>%
  summarize(total_claim_paid = sum(claim_paid, na.rm = TRUE), .groups = 'drop')

# Step 2: Join the summary back to test_data_prem
test_data_prem <- test_data_prem %>%
  left_join(claims_summary, by = "exposure_id")

# Replace NA values in total_claim_paid with 0
test_data_prem$total_claim_paid[is.na(test_data_prem$total_claim_paid)] <- 0

# Optionally, check the changes
print(head(test_data_prem$total_claim_paid))  # Display the first few entries

# Assuming previous code remains the same

# After calculating total_result and ensuring test_data_prem is correctly set up
test_data_prem$total_claim_paid[is.na(test_data_prem$total_claim_paid)] <- 0

# Ensure that both vectors are the same length
if(length(test_data_prem$total_claim_paid) == length(total_result)) {
  # Calculate the pairwise squared differences
  squared_differences <- (test_data_prem$total_claim_paid - total_result) ^ 2
  
  # Sum all squared differences
  total_squared_difference <- sum(squared_differences)
  
  # Print the total squared difference
  print(paste("Total Pairwise Squared Difference:", total_squared_difference))
  
  # Calculate RMSE
  rmse <- sqrt(total_squared_difference / length(test_data_prem$total_claim_paid))
  print(paste("RMSE:", rmse))
} else {
  print("Error: Lengths of total_claim_paid and total_result do not match.")
}
```

### Pricing Output

First, prepare the new customer profiles to match the required format for the model, importing the new datasets and reformatting variables.

```{r}
### Final_pred dataset set up

### Run 'Adding_data2 (new)' sheet first

final_pred_new <- read.csv('New_Customers_Pricing_Output_File.csv')

final_pred <- read.csv('New_Customers_Pricing_Output_File.csv')

final_pred <- final_pred %>%
  filter(!nb_breed_trait %in% c('siamese', 'forest cat', 'spotted cats', 'persian', 'burmese'))

final_pred <- final_pred[ , !names(final_pred) %in% "Full_month_premium"]


final_pred_new <- final_pred_new %>%
  filter(!nb_breed_trait %in% c('siamese', 'forest cat', 'spotted cats', 'persian', 'burmese'))

final_pred_new <- final_pred_new[ , !names(final_pred_new) %in% "Full_month_premium"]


final_pred$person_dob <- as.Date(final_pred$person_dob, format = "%d/%m/%Y")

final_pred$quote_time_group <- as.Date(final_pred$quote_time_group, format = "%d/%m/%Y")

final_pred$pet_gender <- factor(final_pred$pet_gender, levels = c("male", "female"))
final_pred$pet_de_sexed <- factor(final_pred$pet_de_sexed, levels = c("TRUE", "FALSE"))
final_pred$nb_address_type_adj <- factor(final_pred$nb_address_type_adj)
final_pred$is_multi_pet_plan <- factor(final_pred$is_multi_pet_plan, levels = c("TRUE", "FALSE"))

# Summarise the data, accounting for all columns
final_pred <- final_pred %>%
  group_by(exposure_id) %>%
  summarise(
    # Numeric Columns
    pet_age_months = mean(pet_age_months, na.rm = TRUE),  # Average pet age in months
    pet_age_years = mean(pet_age_years, na.rm = TRUE),  # Average pet age in years
    nb_contribution = mean(nb_contribution, na.rm = TRUE), # Average contribution
    nb_contribution_excess = mean(nb_contribution_excess, na.rm = TRUE),  # Average contribution excess
    nb_excess = mean(nb_excess, na.rm = TRUE),  # Average excess
    nb_number_of_breeds = max(nb_number_of_breeds, na.rm = TRUE),  # Maximum number of breeds
    nb_average_breed_size = mean(nb_average_breed_size, na.rm = TRUE),  # Average breed size
    
    # Boolean Columns
    is_multi_pet_plan = any(is_multi_pet_plan == "TRUE"),  # TRUE if any multi-pet plan exists
    pet_de_sexed = first(pet_de_sexed),  # First occurrence of pet de-sexed
    
    # Categorical Columns
    pet_gender = first(pet_gender),  # First occurrence of pet gender
    pet_de_sexed_age = first(pet_de_sexed_age),  # First occurrence of pet de-sexed age
    nb_address_type_adj = first(nb_address_type_adj),  # First occurrence of address type
    nb_postcode = first(nb_postcode),  # First postcode
    nb_state = first(nb_state),  # First state
    nb_suburb = first(nb_suburb),  # First suburb
    nb_breed_type = first(nb_breed_type),  # First breed type
    nb_breed_trait = first(nb_breed_trait),  # First breed trait
    nb_breed_name_unique = first(nb_breed_name_unique),  # First breed name
    nb_breed_name_unique_concat = first(nb_breed_name_unique_concat),  # First concatenated breed name
    quote_time_group = first(quote_time_group),  # First quote time group
    
    # Person-related columns (if applicable)
    person_dob = first(person_dob),  # First person date of birth
    owner_age_years = max(owner_age_years, na.rm = TRUE),  # Average owner age
    
    # Additional Columns
    pet_is_switcher = first(pet_is_switcher),  # First occurrence of switcher
    lead_date_day = first(lead_date_day),  # First occurrence of lead date day
    quote_date = first(quote_date),  # First quote date
  )

# Perform the left join to add SA3_CODE_2011
final_pred <- final_pred %>%
  left_join(SA3_unique, by = c("nb_postcode" = "POSTCODE"))

final_pred <- final_pred %>%
  left_join(income_sa3 %>% select(Code, Median_employee_income_SA3, Mean_employee_income_SA3), 
            by = c("SA3_CODE_2011" = "Code"))

final_pred <- final_pred %>%
  left_join(
    added_var_abs %>% 
      select(Code, 
             Total_number_bus, 
             Number_non.employing_bus,  # Updated to the correct column name
             Total_employed_aged_15_and_over, 
             Median_price_established_house_transfers, 
             Value_private_sector_houses, 
             Agriculture_forestry_fishing_per, 
             Health_care_social_assistance_per, 
             Professional_scientific_technical_services_per, 
             Number_employee_jobs_agriculture_forestry_fishing, 
             Number_employee._jobs_health_care,  # Updated to the correct column name
             Total_persons_aged_15_years_over, 
             Persons_who_made_HELP_repayment, 
             Number_employee_jobs_professional_scientific_technical, 
             Graduate_diploma_perc, 
             Bachelor_degree_perc, 
             Completed_year_12_perc, 
             Worked_from_home, 
             Average_household_size, 
             Lone_person_households, 
             Median_weekly_household_rent, 
             Persons_who_made_gift_or_donation, 
             Total_protected_land_area, 
             National_parks, 
             Nature_reserves, 
             X_1.499_per_week, 
             Participation_rate, 
             Professionals_perc, 
             Bachelor_degree_perc.1, 
             X_3000_or_more_per_week, 
             Managers, 
             Unemployment_rate, 
             X_2000_2999_per_week, 
             Not_in_labour_force), 
    by = c("SA3_CODE_2011" = "Code")
  )

# Convert appropriate columns to factor or numeric
final_pred <- final_pred %>%
  mutate(
    # Convert logical columns to factor
    is_multi_pet_plan = as.factor(is_multi_pet_plan),
    pet_de_sexed = as.factor(pet_de_sexed),
    pet_gender = as.factor(pet_gender),
    pet_de_sexed_age = as.factor(pet_de_sexed_age),
    nb_address_type_adj = as.factor(nb_address_type_adj),
    nb_breed_type = as.factor(nb_breed_type),
    
    # Convert character columns to factor if needed
    nb_state = as.factor(nb_state),
    nb_suburb = as.factor(nb_suburb),
    
    # Ensure numeric variables are correctly encoded
    Median_employee_income_SA3 = as.numeric(as.character(Median_employee_income_SA3)),
    Mean_employee_income_SA3 = as.numeric(as.character(Mean_employee_income_SA3)),
    Total_number_bus = as.numeric(as.character(Total_number_bus)),
    Number_non.employing_bus = as.numeric(as.character(Number_non.employing_bus)),
    Total_employed_aged_15_and_over = as.numeric(as.character(Total_employed_aged_15_and_over)),
    Agriculture_forestry_fishing_per = as.numeric(as.character(Agriculture_forestry_fishing_per)),
    Health_care_social_assistance_per = as.numeric(as.character(Health_care_social_assistance_per)),
    Professional_scientific_technical_services_per = as.numeric(as.character(Professional_scientific_technical_services_per)),
    Number_employee_jobs_agriculture_forestry_fishing = as.numeric(as.character(Number_employee_jobs_agriculture_forestry_fishing)),
    Number_employee._jobs_health_care = as.numeric(as.character(Number_employee._jobs_health_care)),
    Worked_from_home = as.numeric(as.character(Worked_from_home)),
    Lone_person_households = as.numeric(as.character(Lone_person_households)),
    X_1.499_per_week = as.numeric(as.character(X_1.499_per_week)),
    Participation_rate = as.numeric(as.character(Participation_rate)),
    Professionals_perc = as.numeric(as.character(Professionals_perc)),
    Bachelor_degree_perc.1 = as.numeric(as.character(Bachelor_degree_perc.1)),
    X_3000_or_more_per_week = as.numeric(as.character(X_3000_or_more_per_week)),
    Managers = as.numeric(as.character(Managers)),
    Unemployment_rate = as.numeric(as.character(Unemployment_rate)),
    X_2000_2999_per_week = as.numeric(as.character(X_2000_2999_per_week)),
    Not_in_labour_force = as.numeric(as.character(Not_in_labour_force)),
    Completed_year_12_perc = as.numeric(as.character(Completed_year_12_perc)),
    Bachelor_degree_perc = as.numeric(as.character(Bachelor_degree_perc)),
    Graduate_diploma_perc = as.numeric(as.character(Graduate_diploma_perc)),
    Number_employee_jobs_professional_scientific_technical = as.numeric(as.character(Number_employee_jobs_professional_scientific_technical)),
    Persons_who_made_HELP_repayment = as.numeric(as.character(Persons_who_made_HELP_repayment)),
    National_parks = as.numeric(as.character(National_parks)),
    Nature_reserves = as.numeric(as.character(Nature_reserves)),
    Total_protected_land_area = as.numeric(as.character(Total_protected_land_area)),
    Persons_who_made_gift_or_donation = as.numeric(as.character(Persons_who_made_gift_or_donation)),
  )

final_pred$nb_breed_name_unique <- tolower(final_pred$nb_breed_name_unique)  # Convert to lowercase
final_pred$nb_breed_name_unique <- trimws(final_pred$nb_breed_name_unique)   # Remove leading/trailing whitespace


unique_new_dog_breeds2 <- unique(new_dog_data$Breed.Name)

unique_frequency_breeds2 <- unique(final_pred$nb_breed_name_unique)

missing_breeds2 <- setdiff(unique_frequency_breeds2, unique_new_dog_breeds2)

distance_matrix2 <- stringdistmatrix(missing_breeds2, unique_new_dog_breeds2, method = "jw")  # Jaro-Winkler method

closest_match_indices2 <- apply(distance_matrix2, 1, which.min)

# Get the closest matches from new_dog_data
closest_matches2 <- unique_new_dog_breeds[closest_match_indices2]

# Create a data frame with missing breeds and their closest match
matching_df2 <- data.frame(
  Missing_Breeds = missing_breeds2,
  Suggested_Matches = closest_matches2,
  stringsAsFactors = FALSE
)

replacement_vector2 <- setNames(matching_df2$Suggested_Matches, matching_df2$Missing_Breeds)

# Replace values in 'nb_breed_name_unique' with the corresponding 'Suggested_Matches' values
final_pred$nb_breed_name_unique <- ifelse(
  final_pred$nb_breed_name_unique %in% matching_df2$Missing_Breeds, 
  replacement_vector2[final_pred$nb_breed_name_unique], 
  final_pred$nb_breed_name_unique
)

# Merge frequency_data with new_dog_data on breed names
final_pred <- merge(final_pred, new_dog_data, by.x = "nb_breed_name_unique", by.y = "Breed.Name", all.x = TRUE)

# Convert specified columns to factors
final_pred$Dog.Breed.Group <- as.factor(final_pred$Dog.Breed.Group)
final_pred$Dog.Size <- as.factor(final_pred$Dog.Size)
final_pred$Height <- as.factor(final_pred$Height)
final_pred$Weight <- as.factor(final_pred$Weight)
final_pred$Life.Span <- as.factor(final_pred$Life.Span)

final_pred <- add_exp_future_life(final_pred)

final_pred <- add_apartment_feature(final_pred)

final_pred <- add_dog_safety_feature(final_pred)

final_pred <- add_expected_aggression(final_pred)

# Assuming frequency_data is your original data frame
final_pred <- final_pred %>%
  select(exp_future_life,
         nb_contribution_excess, nb_contribution, nb_excess,
         pet_age_months, nb_average_breed_size,
         is_multi_pet_plan, pet_de_sexed, pet_gender,
         nb_address_type_adj, nb_breed_type, owner_age_years,
         Total_number_bus, Agriculture_forestry_fishing_per,
         Health_care_social_assistance_per, Completed_year_12_perc,
         Average_household_size, Participation_rate,
         Median_employee_income_SA3, Dog.Breed.Group,
         Sensitivity.Level,dog_safety, happy_in_apartment,
         Tolerates.Cold.Weather, Tolerates.Hot.Weather,
         Kid.Friendly, Friendly.Toward.Strangers,
         Drooling.Potential, Easy.To.Groom, General.Health,
         Potential.For.Weight.Gain, Intelligence,
         Potential.For.Mouthiness, Prey.Drive, expected_aggression,
         Tendency.To.Bark.Or.Howl, Wanderlust.Potential,
         Energy.Level, Intensity)

final_pred <- final_pred[, !(names(final_pred) %in% c("Dog.Breed.Group"))]

# Convert these columns to ordinal variables with 5 levels
final_pred <- final_pred %>%
  mutate(across(all_of(cols_to_change), ~ cut(., breaks = 5, labels = 1:5, ordered_result = TRUE)))

library(dplyr)

# Fill NA values with the mean for numeric columns and the most frequent value for ordinal factors
final_pred <- final_pred %>%
  mutate(across(everything(), ~ if (is.numeric(.)) {
    # Replace NA values with the mean for numeric columns
    replace(., is.na(.), mean(., na.rm = TRUE))
  } else if (is.factor(.) || is.ordered(.)) {
    # Replace NA values with the most frequent value for factors/ordinals
    most_frequent_value <- names(sort(table(.), decreasing = TRUE))[1]
    replace(., is.na(.), most_frequent_value)
  } else {
    .
  }))


# Check for missing values and count them for each column
missing_values_count <- sapply(final_pred, function(x) sum(is.na(x)))

# Display the result
missing_values_count

str(final_pred)

str(train_data)




```

Now, run the model on this dataset to generate predictions.

```{r}
# full model training

View(final_pred)


# full data for training:

# Frequency:

full_data <- rbind(train_data, test_data)

full_data_cat1 <- rbind(train_data_cat1, test_data_cat1)

full_data_cat2 <- rbind(train_data_cat2, test_data_cat2)

full_data_cat3 <- rbind(train_data_cat3, test_data_cat3)

full_data_sev_total <- rbind(full_data_cat1, full_data_cat2, full_data_cat3)

#----Frequency Models----

# Load required libraries
library(caret)
library(reticulate)

# Import the Python interpret library
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict
evaluate_and_predict_ebm <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Print the test_data_prem dataframe to see its structure and values before any adjustments
  print("The final_pred (test_data_prem) dataframe before adjustments:")
  print(head(test_data_prem))  # Prints the first few rows
  print(str(test_data_prem))   # Prints the structure of the dataframe
  
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Check if training data is empty after filtering
  if (nrow(train_data) == 0) {
    stop("Training data is empty after removing NAs or specified variables.")
  }
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% target_var, drop = FALSE]
  y_train <- train_data[[target_var]]
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns, drop = FALSE]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, filling with NA
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA
  }
  
  # Handle missing values in test_data_prem_adjusted by filling with 0 (or use other strategies)
  test_data_prem_adjusted[is.na(test_data_prem_adjusted)] <- 0
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var, drop = FALSE]
  
  # Convert data frames to matrices for Python
  X_train_matrix <- as.matrix(X_train)
  X_test_matrix <- as.matrix(X_test_prem)
  
  # Check the dimensions of the matrices
  print(dim(X_train_matrix))
  print(dim(X_test_matrix))
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()
  ebm$fit(X_train_matrix, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_matrix)
  
  # Check if predictions are generated
  if (length(predictions_prem) == 0) {
    stop("No predictions were generated.")
  }
  
  # Ensure predictions are not negative
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}

# Define the list of variables to remove
vars_to_remove <- c("claim_count", "earned_units", "cat1", "cat2", "cat3", 
                    "claim_rate_cat2", "claim_rate_cat3", "Drooling.Potential",
                    "nb_contribution", "Energy.Level", "policy_age_group", 
                    "is_multi_pet_plan", "pet_de_sexed", "UW_season")

# Call the function to predict for test_data_prem using the EBM model for Cat1
predicted_frequencies_prem_cat1 <- evaluate_and_predict_ebm(
  train_data = full_data, 
  test_data_prem = final_pred, 
  target_var = "claim_rate_cat1", 
  vars_to_remove = vars_to_remove
)

# Print the predictions
nrow(predicted_frequencies_prem_cat1)


# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict for Cat2
evaluate_and_predict_ebm_cat2 <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Print the test_data_prem dataframe to see its structure and values before any adjustments
  print("The final_pred (test_data_prem) dataframe before adjustments for Cat2:")
  print(head(test_data_prem))  # Prints the first few rows
  print(str(test_data_prem))   # Prints the structure of the dataframe
  
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Check if training data is empty after filtering
  if (nrow(train_data) == 0) {
    stop("Training data is empty after removing NAs or specified variables.")
  }
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% target_var, drop = FALSE]
  y_train <- train_data[[target_var]]
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns, drop = FALSE]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, filling with NA
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA
  }
  
  # Handle missing values in test_data_prem_adjusted by filling with 0 (or use other strategies)
  test_data_prem_adjusted[is.na(test_data_prem_adjusted)] <- 0
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var, drop = FALSE]
  
  # Convert data frames to matrices for Python
  X_train_matrix <- as.matrix(X_train)
  X_test_matrix <- as.matrix(X_test_prem)
  
  # Check the dimensions of the matrices
  print(dim(X_train_matrix))
  print(dim(X_test_matrix))
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()
  ebm$fit(X_train_matrix, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_matrix)
  
  # Check if predictions are generated
  if (length(predictions_prem) == 0) {
    stop("No predictions were generated.")
  }
  
  # Ensure predictions are not negative
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}


# Define the list of variables to remove for Cat2 model
vars_to_remove_cat2 <- c("claim_count", "earned_units", "cat1", "cat2", "cat3", 
                         "claim_rate_cat1", "claim_rate_cat3", "is_multi_pet_plan", 
                         "pet_de_sexed", "UW_season", "Energy.Level", "policy_age_group")

# Call the function to predict for test_data_prem using the EBM model for Cat2
predicted_frequencies_prem_cat2 <- evaluate_and_predict_ebm_cat2(train_data = full_data, 
                                                                 test_data_prem = final_pred, 
                                                                 target_var = "claim_rate_cat2", 
                                                                 vars_to_remove = vars_to_remove_cat2)


# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Function to adjust test_data_prem columns, build EBM, and predict for Cat3
evaluate_and_predict_ebm_cat3 <- function(train_data, test_data_prem, target_var, vars_to_remove) {
  # Print the test_data_prem dataframe to see its structure and values before any adjustments
  print("The final_pred (test_data_prem) dataframe before adjustments for Cat3:")
  print(head(test_data_prem))  # Prints the first few rows
  print(str(test_data_prem))   # Prints the structure of the dataframe
  
  # Remove specified variables from training set
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Remove rows with NAs in training data
  train_data <- na.omit(train_data)
  
  # Check if training data is empty after filtering
  if (nrow(train_data) == 0) {
    stop("Training data is empty after removing NAs or specified variables.")
  }
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% target_var, drop = FALSE]
  y_train <- train_data[[target_var]]
  
  # Check for extra columns in test_data_prem that are not in train_data
  extra_columns <- setdiff(colnames(test_data_prem), colnames(train_data))
  
  # Remove extra columns from test_data_prem
  test_data_prem_adjusted <- test_data_prem[, !names(test_data_prem) %in% extra_columns, drop = FALSE]
  
  # Check for missing columns in test_data_prem that are present in train_data
  missing_columns <- setdiff(colnames(train_data), colnames(test_data_prem_adjusted))
  
  # Add missing columns to test_data_prem_adjusted, filling with NA
  for (col in missing_columns) {
    test_data_prem_adjusted[[col]] <- NA
  }
  
  # Handle missing values in test_data_prem_adjusted by filling with 0 (or use other strategies)
  test_data_prem_adjusted[is.na(test_data_prem_adjusted)] <- 0
  
  # Prepare predictor variables for test_data_prem
  X_test_prem <- test_data_prem_adjusted[, !names(test_data_prem_adjusted) %in% target_var, drop = FALSE]
  
  # Convert data frames to matrices for Python
  X_train_matrix <- as.matrix(X_train)
  X_test_matrix <- as.matrix(X_test_prem)
  
  # Check the dimensions of the matrices
  print(dim(X_train_matrix))
  print(dim(X_test_matrix))
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()
  ebm$fit(X_train_matrix, y_train)
  
  # Make predictions using the adjusted test data prem
  predictions_prem <- ebm$predict(X_test_matrix)
  
  # Check if predictions are generated
  if (length(predictions_prem) == 0) {
    stop("No predictions were generated.")
  }
  
  # Ensure predictions are not negative
  if (any(predictions_prem < 0)) {
    smallest_positive_value <- min(predictions_prem[predictions_prem > 0], na.rm = TRUE)
    predictions_prem[predictions_prem < 0] <- smallest_positive_value
  }
  
  # Return the predictions
  return(predictions_prem)
}

# Define the list of variables to remove for Cat3 model
vars_to_remove_cat3 <- c("cat1", "cat2", "cat3", "earned_units", "claim_count", "claim_rate_cat2", 
                         "claim_rate_cat1", "UW_season", "is_multi_pet_plan", "policy_age_group", 
                         "Energy.Level", "Intelligence", "Kid.Friendly", "pet_gender", "Sensitivity.Level")

# Call the function to predict for test_data_prem using the EBM model for Cat3
predicted_frequencies_prem_cat3 <- evaluate_and_predict_ebm_cat3(train_data = full_data, 
                                                                 test_data_prem = final_pred, 
                                                                 target_var = "claim_rate_cat3", 
                                                                 vars_to_remove = vars_to_remove_cat3)
# potentially could fit glm for model 1 frequency

#----Severity models----

# Load necessary libraries
library(reticulate)

# Load the interpret library from Python
interpret <- import("interpret")
interpret_glassbox <- interpret$glassbox

# Variables to remove from the model
vars_to_remove <- c("Sensitivity.Level", "Kid.Friendly", "Drooling.Potential", "Potential.For.Mouthiness", "Avg..Life.Span..years", "Friendly.Toward.Strangers",
                    "Energy.Level", "Easy.To.Groom", "Dog.Friendly", "Intelligence", "Prey.Drive", "pet_gender", "Median_employee_income_SA3", "Intensity", "Tolerates.Hot.Weather",
                    "nb_address_type_adj", "inception_month", "owner_age_years", "Total_number_bus", "policy_age_group", "UW_season")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, vars_to_remove) {
  # Remove the specified variables from the training dataset
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster"), drop = FALSE]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat1 <- train_ebm_model(full_data_cat1, vars_to_remove)

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat1, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem1 <- predict_claim_paid(ebm_model = ebm_model_cat1, 
                                                 test_data_prem = final_pred, 
                                                 train_data_cat1 = full_data_cat1, 
                                                 vars_to_remove = vars_to_remove)



# Variables to remove from the model
vars_to_remove <- c("UW_season", "Intelligence", "Kid.Friendly", "Average_household_size", "nb_address_type_adj", "pet_de_sexed",
                    "Sensitivity.Level", "Friendly.Toward.Strangers", "Tendency.To.Bark.Or.Howl", "Energy.Level", "Prey.Drive", "Easy.To.Groom",
                    "Potential.For.Weight.Gain", "Tolerates.Cold.Weather", "Total_number_bus", "policy_age_group")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, vars_to_remove) {
  # Remove the specified variables from the training dataset
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster"), drop = FALSE]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat2 <- train_ebm_model(full_data_cat2, vars_to_remove)

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat2, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem2 <- predict_claim_paid(ebm_model = ebm_model_cat2, 
                                                 test_data_prem = final_pred, 
                                                 train_data_cat2 = full_data_cat2, 
                                                 vars_to_remove = vars_to_remove)



# Variables to remove from the model
vars_to_remove <- c("Potential.For.Weight.Gain", "Drooling.Potential", "Kid.Friendly", "owner_age_years", "Intelligence", "pet_gender",
                    "nb_address_type_adj", "UW_season", "Sensitivity.Level", "policy_age_group")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, vars_to_remove) {
  # Remove the specified variables from the training dataset
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster"), drop = FALSE]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_cat3 <- train_ebm_model(train_data_cat3, vars_to_remove)


# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat3, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_prem3 <- predict_claim_paid(ebm_model = ebm_model_cat3, 
                                                 test_data_prem = final_pred, 
                                                 train_data_cat3 = full_data_cat3, 
                                                 vars_to_remove = vars_to_remove)

#----Total severity model----
# Variables to remove from the model
vars_to_remove <- c("UW_season", "policy_age_group")

# Function to build and fit the EBM model without specific variables
train_ebm_model <- function(train_data, vars_to_remove) {
  # Remove the specified variables from the training dataset
  train_data <- train_data[, !names(train_data) %in% vars_to_remove]
  
  # Prepare predictor variables and target variable for training
  X_train <- train_data[, !names(train_data) %in% c("claim_paid", "merged_cluster"), drop = FALSE]  # Predictor variables for training
  y_train <- train_data$claim_paid  # Target variable for training
  
  # Create and train the EBM model
  ebm <- interpret_glassbox$ExplainableBoostingRegressor()  # For regression
  ebm$fit(X_train, y_train)
  
  return(ebm)
}

# Train the EBM model without the specified variables
ebm_model_total <- train_ebm_model(full_data_sev_total, vars_to_remove)

### Predict on test_data_prem

# Function to ensure correct columns and make predictions
predict_claim_paid <- function(ebm_model, test_data_prem, train_data_cat2, vars_to_remove) {
  # Remove unnecessary variables from both the training and test datasets
  test_data_prem <- test_data_prem[, !names(test_data_prem) %in% vars_to_remove]
  train_data_cat1 <- train_data_cat1[, !names(train_data_cat1) %in% vars_to_remove]
  
  # Identify the predictor columns used in training (excluding "claim_paid")
  matching_columns <- setdiff(colnames(train_data_cat1), "claim_paid")
  
  # Check for missing columns in test_data_prem
  missing_columns <- setdiff(matching_columns, colnames(test_data_prem))
  if (length(missing_columns) > 0) {
    stop(paste("The following columns are missing from test_data_prem:", paste(missing_columns, collapse = ", ")))
  }
  
  # Check for extra columns in test_data_prem that were not in train_data_cat1
  extra_columns <- setdiff(colnames(test_data_prem), matching_columns)
  if (length(extra_columns) > 0) {
    # Remove the extra columns from test_data_prem
    test_data_prem <- test_data_prem[, !names(test_data_prem) %in% extra_columns]
  }
  
  # Ensure that test_data_prem contains the same predictors as train_data_cat1
  test_data_prem_adjusted <- test_data_prem[, matching_columns, drop = FALSE]
  
  # Make predictions using the adjusted test data
  predicted_claim_paid <- ebm_model$predict(test_data_prem_adjusted)
  
  return(predicted_claim_paid)
}

# Make predictions using the adjusted test_data_prem
predicted_claim_paid_total <- predict_claim_paid(ebm_model = ebm_model_total, 
                                                 test_data_prem = final_pred, 
                                                 train_data_cat2 = full_data_sev_total, 
                                                 vars_to_remove = vars_to_remove)

#----Credibility Adjustment----

## Bulmann-Straub Approach
# Function to calculate scaled credibility factors based on the number of rows in the training data
calculate_scaled_credibility <- function(train_data, overall_data, response_variable, total_rows) {
  # Number of observations in the category
  n <- nrow(train_data)
  
  # Mean of the response variable for individual model
  mean_individual <- mean(train_data[[response_variable]], na.rm = TRUE)
  
  # Overall mean of the response variable
  mean_overall <- mean(overall_data[[response_variable]], na.rm = TRUE)
  
  # Within-group variance (individual risk category)
  var_within <- var(train_data[[response_variable]], na.rm = TRUE)
  
  # Between-group variance (variance between individual mean and overall mean)
  var_between <- mean((mean_individual - mean_overall)^2)
  
  # Unscaled credibility factor Z (traditional formula)
  Z <- n / (n + (var_within / var_between))
  
  # Scale the credibility factor by the number of rows relative to the total
  Z_scaled <- Z
  
  return(list(credibility_factor = Z_scaled, mean_individual = mean_individual, mean_overall = mean_overall))
}

# Total number of rows in the overall data
total_rows <- nrow(train_data_total)

# Example calculation for train_data_cat1, train_data_cat2, train_data_cat3
credibility_cat1 <- calculate_scaled_credibility(train_data_cat1, train_data_total, "claim_paid", total_rows)
credibility_cat2 <- calculate_scaled_credibility(train_data_cat2, train_data_total, "claim_paid", total_rows)
credibility_cat3 <- calculate_scaled_credibility(train_data_cat3, train_data_total, "claim_paid", total_rows)

# Scaled Credibility factors for each category
Z_cat1 <- credibility_cat1$credibility_factor
Z_cat2 <- credibility_cat2$credibility_factor
Z_cat3 <- credibility_cat3$credibility_factor

# Now let's calculate the adjusted severity for each category
# Assuming you have predicted severities from the individual models and the overall model

# Adjusted severity calculation using the scaled credibility factors
adjusted_severity_cat1 <- Z_cat1 * predicted_claim_paid_prem1 + (1 - Z_cat1) * predicted_claim_paid_total
adjusted_severity_cat2 <- Z_cat2 * predicted_claim_paid_prem2 + (1 - Z_cat2) * predicted_claim_paid_total
adjusted_severity_cat3 <- Z_cat3 * predicted_claim_paid_prem3 + (1 - Z_cat3) * predicted_claim_paid_total


predicted_frequencies_prem_cat1
predicted_frequencies_prem_cat2
predicted_frequencies_prem_cat3

adjusted_severity_cat1
adjusted_severity_cat2
adjusted_severity_cat3

# Multiply corresponding entries
result_cat1 <- predicted_frequencies_prem_cat1 * adjusted_severity_cat1
result_cat2 <- predicted_frequencies_prem_cat2 * adjusted_severity_cat2
result_cat3 <- predicted_frequencies_prem_cat3 * adjusted_severity_cat3

# Perform the element-wise sum
total_result <- result_cat1 + result_cat2 + result_cat3

Full_month_premium <- total_result

# Assuming total_result and final_pred are both vectors/dataframes of the same length
final_pred_new$Full_month_premium <- Full_month_premium

# Export the updated dataframe as a CSV file
write.csv(final_pred_new, "final_pred_with_premiums.csv", row.names = FALSE)

# This will create a file named 'final_pred_with_premiums.csv' in your working directory


colnames(final_pred_new)

set.seed(123)
# Create a single train-test split for the frequency data (80% train, 20% test)
train_index <- sample(seq_len(nrow(frequency_data)), size = 0.8 * nrow(frequency_data))
train_data <- frequency_data[train_index, ]
test_data <- frequency_data[-train_index, ]

test_data_prem <- frequency_data2[-train_index, ]

# Step 1: Summarize claim_paid in the claims dataset
claims_summary <- claims %>%
  group_by(exposure_id) %>%
  summarize(total_claim_paid = sum(claim_paid, na.rm = TRUE), .groups = 'drop')

# Step 2: Join the summary back to test_data_prem
test_data_prem <- test_data_prem %>%
  left_join(claims_summary, by = "exposure_id")

# Replace NA values in total_claim_paid with 0
test_data_prem$total_claim_paid[is.na(test_data_prem$total_claim_paid)] <- 0

# Optionally, check the changes
print(head(test_data_prem$total_claim_paid))  # Display the first few entries

# Assuming previous code remains the same

# After calculating total_result and ensuring test_data_prem is correctly set up
test_data_prem$total_claim_paid[is.na(test_data_prem$total_claim_paid)] <- 0

# Ensure that both vectors are the same length
if(length(test_data_prem$total_claim_paid) == length(total_result)) {
  # Calculate the pairwise squared differences
  squared_differences <- (test_data_prem$total_claim_paid - total_result) ^ 2
  
  # Sum all squared differences
  total_squared_difference <- sum(squared_differences)
  
  # Print the total squared difference
  print(paste("Total Pairwise Squared Difference:", total_squared_difference))
  
  # Calculate RMSE
  rmse <- sqrt(total_squared_difference / length(test_data_prem$total_claim_paid))
  print(paste("RMSE:", rmse))
} else {
  print("Error: Lengths of total_claim_paid and total_result do not match.")
}
```
